Supporting Information

### 2016-09-14 portfolio/text/ms.Rmd
Develop methods
Also moved some of the methods to the main paper

Individuals relying on natural resources as their primary source of income
often experience high levels of income variability
due to the combined effects of environmental, biological, and price variation.
Commercial fishers are a prime example and
recent work has sought to identify regulations and fishing behaviours
that can reduce this variability.
One candidate has been that fishing higher levels of species diversity
decreases income variability.
However, it is unknown if this species-diversity effect
on income variability holds
across multiple approaches to diversifying catch,
at the level of individual permit holders,
and across different fishing strategies.
Here, we simultaneously estimate the effects of three
methods of diversification for Alaskan fishery permit holders
on revenue and revenue variability:
(1) owning multiple permits,
(2) owning permits that inherently fish higher species diversity,
and (3) increasing species diversity within the constraints of a permit set.
We control for changes in effort and other
latent effects through time.
We find that
fishing an additional permit is associated with higher revenue
and lower variability (e.g., 
`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
expected standard deviation
for halibut plus another permit vs. single halibut permits); 
fishing permits with higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit 
increase in mean Simpson's diversity index);
but that increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability) 
associated with increased revenue variability.
At the same time, Alaskan commercial fisheries
have been increasingly focussed on
single-permit (often salmon) fishing strategies
with single-permit fishers earning
more than multiple-permit fishers
for the first time in 2013.
Our results suggest that the benefit of species diversity
for revenue variability
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification ---
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

@@

Despite this existing work,
key questions remain on the relationship between species diversity and
fishing revenue variability.
First, we do not know the effect of changing species diversity
for individual permit holders on a year-by-year basis ---
existing approaches have worked at the vessel or community level
and condensed time series into single metrics of variability or risk.
Second, we do not know the effect of changing species diversity
on revenue itself ---
existing approaches have focused solely on variability or risk.
Third, we do not know if the approach to
diversifying affects the relationship between
species diversity and revenue variability.
A fisheries permit holder can purchase
an additional permit to target other species,
pick a permit that inherently targets many species,
or increase species diversity
within the constraints of his or her existing permits.
Many permit holders may be limited to the last approach
given that the first two may be costly and limited by regulations,
but it is unclear if the benefits and costs are the same.
Here, we address these questions by
simultaneously estimating the effects of
these three approaches to species diversification
for Alaskan fishery permit holders on
annual revenue and revenue variability.
Our model makes use of the entire time series,
controls for effort and latent time effects,
and allows the effect of species diversity
to vary by fishing strategy (sets of similar permits)
in a hierarchical model.

@@

[@standevelopmentteam2015]
[@rcoreteam2016]
[@simpson1949]

>>>>

Individuals relying on natural resources as their primary source of income
often experience high levels of income variability
due to the combined effects of environmental, biological, and price variation.
Commercial fishers are a prime example and
recent work has sought to identify regulations and fishing behaviours
that can reduce this variability.
One candidate has been that fishing higher levels of species diversity
decreases income variability.
However, it is unknown if this species-diversity effect
on income variability holds
across multiple approaches to diversifying catch,
at the level of individual permit holders,
and across different fishing strategies.
Here, we simultaneously estimate the effects of three
methods of diversification for Alaskan fishery permit holders
on revenue and revenue variability:
(1) owning multiple permits,
(2) owning permits that inherently fish higher species diversity,
and (3) increasing species diversity within the constraints of a permit set.
We control for changes in effort and other
latent effects through time.
We find that
fishing an additional permit is associated with higher revenue
and lower variability (e.g., 
`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
expected standard deviation
for halibut plus another permit vs. single halibut permits); 
fishing permits with higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit 
increase in mean effective species diversity);
but that increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability) 
associated with increased revenue variability.
At the same time, Alaskan commercial fisheries
have been increasingly focussed on
single-permit (often salmon) fishing strategies
with single-permit fishers earning
more than multiple-permit fishers
for the first time in 2013.
Our results suggest that the benefit of species diversity
for revenue variability
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification ---
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

@@

Despite this existing work,
key questions remain on the relationship between species diversity and
fishing revenue variability.
First, the effect of changing species diversity
for individual permit holders on a year-by-year basis is unknown ---
existing approaches have worked at the vessel or community level
and condensed time series into single metrics of variability or risk.
Second, the effect of changing species diversity
on revenue itself is unknown ---
existing approaches have focused solely on variability or risk.
Third, it is unknown if the approach to
diversifying species fished affects the relationship between
species diversity and revenue variability ---
a fisheries permit holder can purchase
an additional permit to target other species,
pick a permit that inherently targets many species,
or increase species diversity
within the constraints of his or her existing permits.
Many permit holders may be limited to the last approach
given that the first two may be costly and limited by regulations,
but it is unclear if the benefits and costs are the same.
Here, we address these questions by
simultaneously estimating the effects of
these three approaches to species diversification
for Alaskan fishery permit holders on
annual revenue and revenue variability.
Our model makes use of the entire time series,
controls for effort and latent time effects,
and allows the effect of species diversity
to vary by fishing strategy (sets of similar permits)
in a hierarchical model.

@@

- overview

- describe predictors

- We adjusted for inflation by setting all revenues in 2009 USD 
based on data from TODO.

We calculated effective species diversity for
permit holder $i$ and year $t$ as the inverse of 
Simpson's diversity index $\lambda$ [@simpson1949]
weighted by revenue $R$ (herein 'species diversity'):

\begin{equation}
1/\lambda_{i,t} = 1 / \left(\sum\limits_{s=1}^{n_{i,t}}
  \left(R_{i,s,t}/ \sum\limits_{s=1}^{n_{i,t}} R_{i,s,t}\right)\right)^2,
\end{equation}

\noindent
where $s$ indexes species from $1$ through $n$.

The basic model is a hierarchical linear regression with Gaussian errors
predicting $\log$ revenue, $\log(R_{i,t})$, for
each fisher $i$ and time (year) $t$:

\begin{equation}
\log(R_{i,t}) \sim \mathrm{N} \left(\widehat{\log(R_{i,t})}, \widehat{\sigma_{i,t}^2} \right).
\end{equation}

\noindent
Each year $t$ of revenue for fisher $i$
is assigned a strategy indexed by $j$ and
expected $\log(R_{i,t})$ is modelled as:

\begin{equation}
\label{eq:mean}
\widehat{\log(R_{i,t})} = \beta_{0,j,t} -
  \beta_{1,j} S_{i,t} I_{i,t} +
  \beta_{2,j} S_{i,t} I_{i,t} +
  \beta_3 D_{i,t} -
  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
  \log(R_{i,t-1})
\end{equation}

\noindent
where $S_{i,t}$ is the percent change in species diversity,
$\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
$D_{i,t}$ is the percent change in days fished,
$\log \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and
$I_{i,t}$ is an indicator variable that takes a value $0$
if $\mathrm{div}_{i,t} \ge \mathrm{div}_{i,t-1}$
(species diversity increasing; generalizing)
and takes a value $1$ if
$\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$
(species diversity decreasing; specializing).

The term $\log(R_{i,t-1}$) is treated as an offset term:
it is added on the right side of the equation
to essentially model percent change in revenue,
i.e., $\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})$.
While this approach required filtering the data 
to remove person-year combinations without a subsequent year of revenue,
it has a number of advantages: 
it eliminates having to deal with the nonstationary 
timeseries and eliminates having to 
estimate thousands of people-specific intercepts.

The $\beta$'s represent estimated coefficients, some of which are
allowed to vary.
The intercepts $\beta_{0,j,t}$ are allowed to vary by
strategy-year combinations,
are centered on zero, and are constrained by a normal distribution:

\begin{equation}
\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
\end{equation}

\noindent
This, combined with the lack of a global intercept, constrains
the mean change in log revenue across all fishers and years to
be zero. I.e., we assume that the data are stationary
after accounting for all covariates and including
an intercept that varies with strategy-year combinations.
Initial versions of the model
confirmed that the global intercept along with stragey-level
intercepts would be estimated at almost exactly zero.

The slopes $\beta_{1,j}$ and $\beta_{1,j}$
are then allowed to vary by strategy with estimated means:

\begin{align}
\beta_{1,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_1}\right),\\
\beta_{2,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_2}\right).
\end{align}

\noindent
The $\beta_{1,j}$'s represent how much an X\% decrease in species
diversity within a strategy translates to a Y\% change in revenue.
The $\beta_{2,j}$'s represent how much an X\% increase in species
diversity within a strategy translates to a Y\% change in revenue.
These are after controlling for percent change in days fished, $\beta_3$,
and the interaction between days fished and percent change in species
diversity, $\beta_4$ and $\beta_5$.

Whereas a typical hierarchical linear regression model
would estimate a single residual error scale term, $\sigma$,
we model the scale of the residual error  with
another hierarchical model with a similar form:

\begin{equation}
\label{eq:cv}
\widehat{\sigma_{i,t}} =
\exp \left( \gamma_{0,j} -
\gamma_{1,j} S_{i,t} I_{i,t} +
\gamma_{2,j} S_{i,t} I_{i,t} +
\gamma_3 D_{i,t} -
\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
\end{equation}

\noindent
Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively,
percent change in species
diversity and an indicator variable for increasing or
decreasing species diversity.
We exponentiate the equation to ensure all scale parameters, $\sigma_{i,t}$,
are positive. The $\gamma$'s are estimated coefficients, three of which
are allowed to vary by strategy, constrained by normal distributions:

\begin{align}
\gamma_{0,j}  &\sim \mathrm{N} \left(\gamma_0 + \eta_1 M_j + \eta_2 E_j, \sigma^2_{\gamma_0}\right),\\
\gamma_{1,j}  &\sim \mathrm{N} \left(\gamma_1, \sigma^2_{\gamma_1}\right),\\
\gamma_{2,j}  &\sim \mathrm{N} \left(\gamma_2, \sigma^2_{\gamma_2}\right).
\end{align}

\noindent
The mean of the $\gamma_{0,j}$ coefficients is
modelled with
two strategy-level predictors.
The symbols $M_j$ and $E_j$ represent the raw mean species diversity and
mean combined fishing season length (effort) for strategy $j$.
The coefficients $\eta_1$ and $\eta_2$ represent predictors
in this strategy-level regression
estimating effects across strategies.

We fit our models with Stan 2.11.1
[@hoffman2014; @standevelopmentteam2015; @standevelopmentteam2015a; @carpenter2015]
and R 3.3.1 [@rcoreteam2016].
Stan implements the 
No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm [@hoffman2014]
to perform Bayesian statistical inference.
We ran four chains and 2,000 iterations and discarded
the first 1,000 iterations of each chain
as warm up (i.e. 4,000 final samples). 
We checked for chain convergence visually with trace plots.
We also ensured that $\hat{R}<1.05$ (the potential scale reduction factor)
and that the effective sample size was 
greater than 200 for all parameters [@gelman2014].
Code to reproduce our results (with a simulated data set)
is available at <https://github.com/NCEAS/pfx-commercial>.

### 2016-09-14 portfolio/text/supp.Rnw
Strip white space and space out supplement
Develop methods
Also moved some of the methods to the main paper

@@

% Rows of data: \Sexpr{ndat$final}
% Number of permit holders: \Sexpr{npholders}
% Range of years: \Sexpr{range_yrs}
% Number of strategy-years: \Sexpr{nstr_yrs}
% \Sexpr{nstrat$enough_pholders}
% \Sexpr{round(ndat$rev_enough_pholders/1e9,2)}
% lost \% rev no strat:
% \Sexpr{round(ndat$rev_after_no_na_strat/ndat$rev_b4_no_na_strat,2)*100}
% lost rows no strat:
% \Sexpr{round(ndat$after_no_na_strat/ndat$b4_no_na_strat,2)*100}

@@

Describe predictors

We adjusted for inflation by setting all revenues 2009 USD.

We calculated Simpson's diversity index weighted by revenue
as: 1/sum((x/sum(x)) squared)

Overview... why first diff, why 2 part

Negatives are that it uses a smaller subset of the data;
positives are
that it doesn't require estimating people-specific random effects (n = 3000+)
and removes time effects.

The basic model is a hierarchical linear regression with Gaussian errors
predicting $\log$ revenue, $\log(R_{i,t})$, for
each fisher $i$ and time (year) $t$:

\begin{equation}
\log(R_{i,t}) \sim \mathrm{N} \left(\widehat{\log(R_{i,t})}, \widehat{\sigma_{i,t}^2} \right).
\end{equation}

Each year $t$ of revenue for fisher $i$
is assigned a strategy indexed by $j$ and
expected $\log(R_{i,t})$ is modelled as:

\begin{equation}
\label{eq:mean}
\widehat{\log(R_{i,t})} = \beta_{0,j,t} -
  \beta_{1,j} S_{i,t} I_{i,t} +
  \beta_{2,j} S_{i,t} I_{i,t} +
  \beta_3 D_{i,t} -
  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
  \log(R_{i,t-1})
\end{equation}

\noindent
where $S_{i,t}$ is the percent change in species diversity,
$\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
$D_{i,t}$ is the percent change in days fished,
$\log \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and
$I_{i,t}$ is an indicator variable that takes a value $0$
if $\mathrm{div}_{i,t} \ge \mathrm{div}_{i,t-1}$
(species diversity increasing; generalizing)
and takes a value $1$ if
$\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$
(species diversity decreasing; specializing).
$\log(R_{i,t-1}$) is treated as an offset term:
it is added on the right side of the equation
to essentially model percent change in revenue,
i.e., $\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})$.

The $\beta$'s represent estimated coefficients, some of which are
allowed to vary.
The intercepts $\beta_{0,j,t}$ are allowed to vary by
strategy-year combinations,
are centered on zero, and are constrained by a normal distribution:

\begin{equation}
\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
\end{equation}

This, combined with the lack of a global intercept, constrains
the mean change in log revenue across all fishers and years to
be zero. I.e., we assume that the data are stationary
after accounting for all covariates and including
an intercept that varies with strategy-year combinations.
Initial versions of the model
confirmed that the global intercept along with stragey-level
intercepts would be estimated at almost exactly zero.

TODO EXPLAIN WHY DO THIS

The slopes $\beta_{1,j}$ and $\beta_{1,j}$
are then allowed to vary by strategy with estimated means:

\begin{align}
\beta_{1,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_1}\right),\\
\beta_{2,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_2}\right).
\end{align}

The $\beta_{1,j}$'s represent how much an X\% decrease in species
diversity within a strategy translates to a Y\% change in revenue.
The $\beta_{2,j}$'s represent how much an X\% increase in species
diversity within a strategy translates to a Y\% change in revenue.
These are after controlling for percent change in days fished, $\beta_3$,
and the interaction between days fished and percent change in species
diversity, $\beta_4$ and $\beta_5$.

Whereas a typical hierarchical linear regression model
would estimate a single residual error scale term, $\sigma$,
we model the scale of the residual error  with
another hierarchical model with a similar form:

\begin{equation}
\label{eq:cv}
\widehat{\sigma_{i,t}} =
\exp \left( \gamma_{0,j} -
\gamma_{1,j} S_{i,t} I_{i,t} +
\gamma_{2,j} S_{i,t} I_{i,t} +
\gamma_3 D_{i,t} -
\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
\end{equation}

Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively,
percent change in species
diversity and an indicator variable for increasing or
decreasing species diversity.
We exponentiate the equation to ensure all scale parameters, $\sigma_{i,t}$,
are positive.

The $\gamma$'s are estimated coefficients, three of which
are allowed to vary by strategy, constrained by normal distributions:

\begin{align}
\gamma_{0,j}  &\sim \mathrm{N} \left(\gamma_0 + \eta_1 M_j + \eta_2 E_j, \sigma^2_{\gamma_0}\right),\\
\gamma_{1,j}  &\sim \mathrm{N} \left(\gamma_1, \sigma^2_{\gamma_1}\right),\\
\gamma_{2,j}  &\sim \mathrm{N} \left(\gamma_2, \sigma^2_{\gamma_2}\right).
\end{align}

The mean of the $\gamma_{0,j}$ coefficients is
modelled with
two strategy-level predictors.
The symbols $M_j$ and $E_j$ represent the raw mean species diversity and
mean combined fishing season length (effort) for strategy $j$.
The coefficients $\eta_1$ and $\eta_2$ represent predictors
in this strategy-level regression.
The effect of $\eta_1$ is shown in Fig.~3.

We fit our models in Stan version 2.11.1. \ldots
\citep{hoffman2014, standevelopmentteam2015a, carpenter2015a, standevelopmentteam2015}

\citep{gelman2014}
\citep{rcoreteam2016}

@@

>>>>

\linenumbers

\setlength{\parindent}{0.9cm}
\setlength\parskip{0in}

\begin{spacing}{1.3}

@@

TODO length(unique(pholders))
% Rows of data: \Sexpr{ndat$final}
% Number of permit holders: \Sexpr{npholders}
% Range of years: \Sexpr{range_yrs}
% Number of strategy-years: \Sexpr{nstr_yrs}
% \Sexpr{nstrat$enough_pholders}
% \Sexpr{round(ndat$rev_enough_pholders/1e9,2)}
% lost \% rev no strat:
% \Sexpr{round(ndat$rev_after_no_na_strat/ndat$rev_b4_no_na_strat,2)*100}
% lost rows no strat:
% \Sexpr{round(ndat$after_no_na_strat/ndat$b4_no_na_strat,2)*100}

@@

@@

\end{spacing}
